---
title: "DHH2021"
subtitle: "Personal portfolio"
author: "Joni Oksanen"
date: "`r format(Sys.Date(), '%d.%m.%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<!--
1)  Discuss with (or write an email to) your group leader to get information what extra work you need to do to earn the 5 credits.
2)  Choose a public venue (blog post, google docs etc.) for your portfolio.
3)  Write a documentation of your work during the hackathon (what you did as part of the group) that tells also about your experience of the project (reflect about what you brought to the group, what new did you learn, what you should learn in the future etc.). Length approximately 750-1000 words.
4)  Collect all the relevant documents, information about your groups work as a sensibly documented collection of links (code, presentation, poster) into your portfolio.
5)  Feel free to add whatever comes to mind into your hackathon portfolio.
-->

# DHH 2021 reflection

  - our [GitHub](https://github.com/dhh21/semparl)
  - our [final blog post at DHH21 blog](https://dhhackathon.wordpress.com/2021/05/28/semparl-cities-in-parliament/)
  - our [Shiny app](https://jonioks.shinyapps.io/citiesinparl/) 


I participated in the Digital Humanities Hackathon (DHH) 2021 as a team member in the Semantic parlament (Semparl) group. For the purposes of this hackathon, we were given access to [ParliamentSampo](https://seco.cs.aalto.fi/projects/semparl/en/), a work-in-progress search engine, where users can search for results from open data provided by the Finnish Parliament. In addition to word searches from text, the data in ParliamentSampo includes automatic background information of each document such as political party information for persons in the data and place names mentioned in the documents. After learning about the ParliamentSampo and how to use it, we started working from the initial topic of “politics and place”. We discussed our interest and backgrounds related to this project, our research theme was set on researching the development of city image in the Finnish parliamentary discussions during economic crises. Our final presentation in the form of a blog post can be read [here](https://dhhackathon.wordpress.com/2021/05/28/semparl-cities-in-parliament/)

My initial interest regarding participation in DHH2021, was to cooperate with people from various backgrounds and to learn and apply methods that I will use in my personal doctoral results, i.e. modern natural language processing and language modeling tools. My background consists of mostly working and studying the quantitative and computational domain. Few years ago, I was interested in working with text data and NLP, so I ended up writing my master’s thesis in statistics about training word embeddings. Before applying to the hackathon, I did not have any previous experience with digital humanities, besides my thesis. However, I was already familiar with the topics that were mentioned in the application form, but have not yet worked with the topics in the context of multidisciplinary research or collaboration.

The first days of the hackathon went so fast as we were trying to find the appropriate framework to conduct our study. After agreeing on the objectives of our brief, but in-depth analysis of parliamentary debates, I aligned myself in the data visualization and manipulation team. Both topics are something I have experience of working with and most of all enjoy spending time with. Even though my personal motivation to join the hackathon was learning something new by doing, I managed to get involved with things I have already done in the past. I did not mind this at all, as I noticed that I am actually good with these things as I immediately had ideas and how to produce them once someone mentioned visualizations. However, this hackathon was a learning event and I could have utilized the chance to do something new.

From a perspective of teamwork and collaborating within our team, I also thought that I would be the most productive by focusing on creating the final outputs (i.e. [plots](https://github.com/dhh21/semparl/blob/main/plots/financial_crisis_rate.gif), [Shiny applications](https://jonioks.shinyapps.io/citiesinparl/)) as there were not that many other team members with a quantitative background. I also did not want to take full control of the whole visualization part and offered the chance for others to be able to join or take lead in the data visualization process.

In general, participating in this 10 day hackathon was an insightful experience. Most of all, I enjoyed the discussions that we had about the research theme. It is not easy to find a common ground without someone making tradeoffs in multidisciplinary research. I learned from my team members how they approached data from the perspective of their own background and discipline and how their thinking process formulated the research questions that they wanted to examine and how their research questions would align with the interests of other team members. In addition, I believe I was able to teach R and also speak about what R is capable of (as most of our team had used mostly Python). I still had to cut down some parts from our final presentation as we simply did neither have the time to implement them completely nor present the results of these analyses during our presentation. 

I also learned that it is important to take into account the time at your disposal. In this sense, the hackathon did not provide realistic settings as usually you don’t have only 10 days to plan _and_ conduct your research. Taking into account the 10 days of hackathon, I think our topic was a bit too ambitious, but nevertheless we managed to deliver results, which I could have not expected to be produced from scratch within 10 days. I believe that our work was showcasing the possibilities of such an extensive linked data set by focusing on something else than examining the traditional party politics or policies. Among these reasons, I was inspired to continue some of the quantitative analyses that we had no time to complete during the hackathon, but we did produce data sets, which can be used in these analyses.

# Further analysis: Examining party politics and regional policies of electoral disctricts of MPs


## Limitations of data

<!--
1. Limitations of data
  - only the cities included in the original DHH-SEMPARL study are considered (-> requires futher analyzing and processing the data to have more reliable and valid results of the RQ/H)

2. Electoral districts
  - plenary session data includes historical electoral districts (-> manually fixed) (2-> create a csv containing all the historical values)
  
3. Parlamentary groups
  - are defined by the MP's last/latest group, which causes some groups to appear before their inauguration as the MP has been member of multiple parlamentary groups in the past.
--->

### City name mentions

Following analyses were conducted using a dataset produced originally as a part of original DHH2021 research project "Cities in parliament". Our previous study only examined 107 munincipalities, which have the [legal status of a city](https://en.wikipedia.org/wiki/List_of_cities_and_towns_in_Finland) as of 2021. There almost certainly is not included mentions, as the electoral district of Lappi includes only one city as the capital city of Helsinki is an electoral district of its own. Thus the results are highly biased towards highly populated areas, which include large cities.


### Electoral distirct

The electoral district were added to cities by using the 2020 division provided by [geofi](https://github.com/rOpenGov/geofi) package, which uses data provided by Statistics Finland. The historical names of each have been harmonized to match their current names, but the municipality level divison has not been taken into account. The some of borders of electoral districts have changed since 1986. The results reflect using the most current division, but they should been changed to reflect the regional division at the moment. However, similar regional divisions were only available from 2012. Having access to historical electoral districts would provide much more reliable results in this context. Also when examining the electoral district of MPs, the information was acquired from ParliamentSampo. If MP was selected from multiple districts, then the preceding information was discarded and the most recent districts was used. This is a limitation in ParliamentSampo biographies, not this analyses.

### Parliamentary groups

If MP has been a member of multiple parliamentary groups during their career, only the latest information was used. Thus it is possible that a parliamentary group can show up in a time period before its inauguration. This information is also acquired from ParliamentSampo and it would be possible to time match the correct parliament group with the date of the plenary speech, but we had no time to implement this during DHH as party politics were not our team projects' the first priority.


### Time span

Our original study covered two 10-year periods: 1) depression in Finland in the 1990's and 2) global financial crisis. This analyses is also separated in to these two periods, which cover 5 years before and after the peak/starting year of each economic crisis. Extending this research setting to cover larger period of time would more likely improve the reliability of the results. This quantitative examination is not really related to these economic crises, but we will use the data sets that were produced, but not analyzed during the hackathon.

## Data

```{r, plots, child = 'plots.Rmd'}
```



## Discussion

We could improve the reliable and validity of the results by fixing the limitations of the data. Fixing the issues related to data is time consuming and labor-intensive and implementing these changes would not really fit in the scope of the hackathon, but fixing them in this research or in ParliamentSampo data would not be impossible. However, this research is more of a novel example of what kind of research could be done related to regional policies and the geography of politics and how the results could be visually represented. 
